Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-05-13 fib2.tex
Worked with Simon today to try to converge to a final version.
--We are going to add a figure 1(d) and another figure to explain the observables / states thing.
--We use a linear length schedule
--Decided to push some qualitative discussion of the algorithm into the figure caption, but haven't done it yet.

%------------------------------------------------------------------------------------------------------------%
\paragraph{Modelling Fibonacci anyons.}

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system. We imagine an idealized Hamiltonian for this 
system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\mathbb{I}}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\mathbb{I}}_i$ the projector to charge $\mathbb{I}$ at site $i$, i.e.~the ground 
space of the model has vacuum (or no anyon) at each site. Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above. We identify this ground space 
as the codespace of our model. Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighbouring sites of $\Lambda$. It was 
seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. Such pair-creation events act 
on neighbouring pairs of sites, and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

Our error-correction algorithm is based on a hard-decision renormalization group (RG) 
decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, forming clusters of nearby anyons and fusing the anyons 
within each cluster, before iterating at increasing length scales.
It terminates when the lattice 
is free of anyons (i.e.~we have returned to the codespace) or the clustering length scale 
exceeds the total size of the lattice.
\cggb{A few more details? What is our length schedule?} 
\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a hard-decision RG decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

\begin{figure*}[th!]
\begin{center}
	{\scriptsize (a)
	\includegraphics[width=0.57\columnwidth]{pair-create.pdf}
	\hskip 4pt
	(b)
	\includegraphics[width=0.57\columnwidth]{syndrome-1.pdf}
	\hskip 4pt
	(c)}
	\includegraphics[width=0.57\columnwidth]{syndrome-2.pdf}
\caption{The simulation of noise processes proceeds by a) creating pairs of anyons on 
neighbouring tiles represented by open curves b) joining the curves within a given tile 
c) measuring the total charge on each tile.}
\label{f:simex}
\end{center}
\end{figure*}

@@

Of course, if we were to consider noise so strong that the clusters percolate over the lattice, 
we would no longer expect to be able to efficiently simulate the system. However, recall that logical 
errors in our system correspond to processes that act over a non-trivial loop of the torus. If 
the noise operations are so common to have percolated, this is an example of a situation 
where we expect error-correction \emph{not} to succeed. Similarly, if the noise has not yet 
percolated by itself, but the clusters are merged into a large (percolated) cluster by the 
decoding algorithm, this is again an example of a process that we expect in general to correspond to a 
failure of error-correction. In this way, we see that the events that we are able to efficiently 
simulate correspond to those for which we expect error-correction to succeed. We 
can simply declare failure on non-simulable events. This allows us to simulate and 
demonstrate the success of error-correction schemes for Fibonacci anyons, without explicitly 
simulating the failure of such schemes.

@@

Our concrete simulation proceeds by associating a \emph{tile} with each site of the 
$L\times L$ periodic square lattice.
In order to keep track of the observables of the system 
we maintain a set of disjoint curves, representing the history of anyons on the lattice.
These curves can also be thought of as a top-down view of the fusion tree:
the anyons comprising the fusion tree are here displayed as points along each curve. \cggb{Is this still true?}
%The ends of curves correspond to leaves of the fusion tree, or anyons. 
In general, the curves wind around the two 
dimensional tiles in a haphazard way determined by the progress of the simulation.
This allows a dynamically generated basis for the fusion space that easily allows for tensor 
decomposition into factors corresponding to disjoint clusters.
For comparison, previous work 
used a single fixed basis for the fusion space (determined by row-major ordering of 
sites)~\cite{Brell2013}.

@@

\begin{enumerate}[i)]
\item The noise process is modelled by a sequence of pair-creation events distributed as 
a Poisson process across each edge. Curves are created crossing the relevant edge 
and terminating on the neighbouring sites.
\item We first join curves that participate in the same tile in order to, 
\cggb{maybe a couple more lines of detail on this and the next point? perhaps also the previous paragraph?}
\item measure the charge on each tile. Any resulting anyon is left somewhere in the tile.
\item The decoder examines this syndrome and clusters nearby anyons. Each cluster 
is then measured (fused), and the decoder continues to cluster and fuse on larger 
scales until there are no more anyons, or a cluster includes a non-trivial loop on the 
lattice. Since operators with support on such non-trivial clusters will generally perform an 
undetectable error on the encoded state, in this case we abandon the simulation and 
declare failure to error correct.
\end{enumerate}

@@

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=\columnwidth]{anyons-kyle.pdf}
\caption{Decoder failure as a function of simulation time for various lattice sizes, showing 
threshold behavior at around $t_{\mathrm{sim}}\simeq 0.125$.}
\label{f:threshold}
\end{center}
\end{figure}

%\cggb{Do we need to define $t_{\mathrm{sim}}$ more explicitly?}

We can guarantee that error-correction will succeed whenever noise does not percolate, but it is possible that percolated events may still result in no error. The connection between the percolation threshold and the error-correction threshold is not well understood in general~\cite{Hastings2014}, though it is clear that our threshold estimate will be a lower bound for the true threshold that may be found if all events (including those that have percolated) were simulated by an optimal decoder. However, by simulating such percolated events on small lattices, our evidence suggests that for the Fibonacci anyon code that we study, the vast majority of such events will result in a logical error. As such it is likely that neglecting such events does not significantly effect our observed threshold value.

@@

We have demonstrated classical simulation of successful error correction in a universal anyon 
model. Though we have chosen several properties of our model and simulation in a 
convenient way for simplicity, we think it very unlikely that they will affect the qualitative 
results, similarly to the results of Ref.~\cite{Brell2013}. In particular, although we have 
modelled our logical qubit as encoded in the global topological degrees of freedom of our 
system, we could have encoded it in the fusion space of several preferred $\tau$ anyons. 
This situation would be appropriate to model error-correction routines for topological 
quantum computation. Additionally, we expect our results to be stable to changes 
in details of the noise model, decoding algorithm, and so on.

>>>>

%------------------------------------------------------------------------------------------------------------%
\paragraph{Modeling Fibonacci anyons.}

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system. We imagine an idealized Hamiltonian for this 
system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\mathbb{I}}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\mathbb{I}}_i$ the projector to charge $\mathbb{I}$ at site $i$, i.e.~the ground 
space of the model has vacuum (or no anyon) at each site. Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above. We identify this ground space 
as the codespace of our model. Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$. It was 
seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. Such pair-creation events act 
on neighboring pairs of sites, and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

Our error-correction algorithm is based on a hard-decision renormalization group (RG) 
decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, forming clusters of nearby anyons and fusing the anyons 
within each cluster, before iterating at increasing length scales.
It terminates when the lattice 
is free of anyons (i.e.~we have returned to the codespace) or the decoder performs a homologically nontrivial operation, thereby altering the codespace.
%\cggb{A few more details? What is our length schedule?} 
%\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a hard-decision RG decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

\begin{figure*}[th!]
\begin{center}
	{\scriptsize (a)
	\includegraphics[width=0.45\columnwidth]{pair-create.pdf}
	\hskip 4pt
	(b)
	\includegraphics[width=0.45\columnwidth]{syndrome-1.pdf}
	\hskip 4pt
	(c)
	\includegraphics[width=0.45\columnwidth]{syndrome-2.pdf}
	\hskip 4pt
	(d)}
	\includegraphics[width=0.45\columnwidth]{syndrome-2.pdf}
\caption{\stf{polish this a bit} The simulation of noise processes proceeds by a) creating pairs of anyons on neighboring tiles represented by open curves b) joining the curves within a given tile c) measuring the total charge on each tile. \stf{Add another graph that shows how we use syndrome info.}}
\label{f:simex}
\end{center}
\end{figure*}

@@

Of course, if we were to consider noise so strong that the clusters percolate over the entire lattice, we would no longer expect to be able to efficiently simulate the system for large $L$. 
However, recall that logical errors in our system correspond to processes that act over a non-trivial loop of the torus. 
Our simulation remains efficient until the point that we start to create nontrivial loops around the torus, and our simulation declares failure at this point, meaning that we can simulate large lattice sizes in this regime, including up to $L=128$.

@@

\stf{Add figure about observables versus states.}

Our concrete simulation proceeds by associating a \emph{tile} with each site of the 
$L\times L$ periodic square lattice.
In order to keep track of the observables of the system 
we maintain a set of disjoint curves, representing the history of anyons on the lattice.
These curves can also be thought of as a top-down view of the fusion tree:
the anyons comprising the fusion tree are here displayed as points along each curve. 
In general, the curves wind around the two 
dimensional tiles in a haphazard way determined by the progress of the simulation.
This allows a dynamically generated basis for the fusion space that easily allows for tensor 
decomposition into factors corresponding to disjoint clusters.
For comparison, previous work 
used a single fixed basis for the fusion space (determined by row-major ordering of 
sites)~\cite{Brell2013}.

@@

\begin{enumerate}[a)]
\item The noise process is modeled by a sequence of pair-creation events distributed as 
a Poisson process across each edge. Curves are created crossing the relevant edge 
and terminating on the neighboring sites.
\item We first join curves that participate in the same tile in order to, 
\cggb{maybe a couple more lines of detail on this and the next point? perhaps also the previous paragraph?}
\item measure the charge on each tile. Any resulting anyon is left somewhere in the tile.
\item The decoder examines this syndrome and clusters nearby anyons. Each cluster 
is then measured (fused), and the decoder continues to cluster and fuse on larger 
scales until there are no more anyons, or a cluster includes a non-trivial loop on the 
lattice. Since operators with support on such non-trivial clusters will generally perform an 
undetectable error on the encoded state, in this case we abandon the simulation and 
declare failure to error correct.
\end{enumerate}

@@

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=\columnwidth]{anyons-kyle.pdf}
\caption{\stf{polish this a bit} Decoder failure as a function of simulation time for various lattice sizes, showing 
threshold behavior at around $t_{\mathrm{sim}}\simeq 0.125$.}
\label{f:threshold}
\end{center}
\end{figure}

We can guarantee that error-correction will succeed whenever noise does not percolate, but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold is not well understood in general~\cite{Hastings2014}, though it is clear that our threshold estimate will be a lower bound for the true threshold that may be found if all events (including those that have percolated) were simulated by an optimal decoder. 
However, calculations of such nontrivial operations suggest that all such processes will indeed result in a logical or leakage error.
As such it is likely that neglecting full simulation for such events does not significantly effect our observed threshold value.

@@

We have demonstrated classical simulation of successful error correction in a universal anyon model. 
Though we have chosen several properties of our model and simulation in a convenient way for simplicity, we think it very unlikely that they will affect the qualitative results, similarly to the results of Ref.~\cite{Brell2013}. 
In particular, although we have modeled our logical qubit as encoded in the global topological degrees of freedom of our system, we could have encoded it in the fusion space of several preferred $\tau$ anyons. 
This situation would be appropriate to model error-correction routines for topological quantum computation. 
Additionally, we expect our results to be stable to changes in details of the noise model, decoding algorithm, and so on.
